---
title: "Oak Data Exploration"
format: html
editor: visual
---

Oak exploratory data analysis and cleaning

```{r}
library(tidyverse)


library(readxl)
Survival_HeightApril26Blocks7_6Clean <- 
  read_excel("Oak Project/Data/Survival_HeightApril26Blocks7&6Clean.xlsx", 
                                sheet = "Input1", skip = 1,
             na = "NA")

#View(Survival_HeightApril26Blocks7_6Clean)

names(Survival_HeightApril26Blocks7_6Clean)
```

Rename columns

```{r}
Survival_HeightApril26Blocks7_6Clean <- 
  
  Survival_HeightApril26Blocks7_6Clean |> 

  rename(Page_Num = "Page #",
         Date = "MM/DD/YYYY",
        Recorder  = "Name",
        Block  = "#...4",
        Tag_Num  = "#...5",
        Oak_type  = "NR, S, or A",
       Treatment   = "AC, OA, NP, DT, or DL",
       OA_microclimate   = "TH, TI, CN",
       Seedling_Ht_cm   =  "(cm)",
       Num_leaves  = "#...10",
       Num_dead_leaves   = "#...11",
        Leaf_greenness  = "%",
       Num_germinated   = "#...13" ,
        RCD  = "(mm)",
       Notes   = "as written on data sheet")

Survival_HeightApril26Blocks7_6Clean
```

```{r}
Survival_HeightApril26Blocks7_6Clean
```

Column Notes:

Oak Type - Oak Type refers to how it was planted if at all.

- A = Acorn
- NR = Natural Recruit
- S = Seedling

Treatments - Treatment refers to microclimate treatment (where it was planted).

- OA = Open Area
- SA = Salvia apiana
- DT = Downed Tree
- AC = Adult canopy
- NP = Nurse plant
- DL = Dripline

Microclimates - OA Microclimate refers to the treatment conducted on open area plots

- CN = Control
- TH = Thatch removed
- TI = Thatch removed AND invasive species removed

Column L (Greenness) is a formula using :=IF(K3=0, 100, 100-(K3/J3)\*100)

if dead leaves = 0, then its 100% green

if not then, 100 - dead leaves/green leaves\*100

- not sure I like this formula! I feel like it ought to be green/dead + green) ? will have to ask Sarah Kimball

```{r}
Survival_HeightApril26Blocks7_6Clean$RCD |> unique()
```

NEEDS:

- Tag nums are confusing - why some with A1, A2, etc?

- seedling height needs to be converted to numeric, figure out what NA means

- num leaves, num dead leaves also to numeric, figure out what NA means

- num germinated, RCD also to numeric

- leaf greenness what is #VALUE in context

- import multiple excel spreadsheets and merge them together

```{r}

```

SOLUTIONS:

Converting rows to numeric is easy by adding an argument to the read_excel() function to interpret "NA" as NA. This works for us because the team that entered data was quite consistent in entering their NAs.

spreadsheet contains both planted acorns and natural recruits. NA is for sites where no planted acorns germinated, therefore most variables not collected about it.

Tag_num actually contains two pieces of information - the site (### and the acorn number (A1, A2, A3).

It would be better if these were separate columns!

There also ought to be rows for each of the planted acorns that did not germinate (e.g., a lot more NAs). That way summaries would make more sense.

to import and merge the spreadsheets, we can use

```{r}
#install.packages("plyr")
library("dplyr")                                                
library("plyr")                                                 
library("readr")  
library("readxl")


gfg_data <- list.files(path = "Oak Project/Data",    
                       pattern = "*.xlsx",
                       full.names = TRUE) |>  
  lapply(read_excel, 
         na = "NA", sheet = "Input1") |> 
  bind_rows                                                      

gfg_data 
```

```{r}
gfg_data |> filter(!Recorder )
```

```{r}
renames <- function(data){
  data <- data |> 
  rename(Page_Num = "Page #",
         Date = "MM/DD/YYYY",
         Recorder  = "Name",
         Block  = "#...4",
         Tag_Num  = "#...5",
         Oak_type  = "NR, S, or A",
         Treatment   = "AC, OA, NP, DT, or DL",
         OA_microclimate   = "TH, TI, CN",
         Seedling_Ht_cm   =  "(cm)",
         Num_leaves  = "#...10",
         Num_dead_leaves   = "#...11",
         Leaf_greenness  = "%",
         Num_germinated   = "#...13" ,
         RCD  = "(mm)",
         Notes   = "as written on data sheet")
  return(data)
  }
```

```{r}
library(tidyverse)

data1 <-  read_excel("Oak Project/Data/Survival_HeightApril26Blocks7&6Clean.xlsx", 
             sheet = "Input1", skip = 1,
             na = "NA") |> 
  renames()


data2 <-   read_excel("Oak Project/Data/Survival_HeightFeb26Clean.xlsx", 
             sheet = "Input1", skip = 1,
             na = "NA")

data3 <-   read_excel("Oak Project/Data/Survival_HeightMarch26Blocks7&6Clean.xlsx", 
             sheet = "Input1", skip = 1,
             na = "NA")

data4 <-  read_excel("Oak Project/Data/Survival_HeightMay26Blocks5-7Clean.xlsx", 
             sheet = "Input1", skip = 1,
             na = "NA")
data5 <- read_excel("Oak Project/Data/Survival_HeightOct2025Clean.xlsx", 
             sheet = "Input1", skip = 1,
             na = "NA")

```
